Investigating social change during cultural contact period using geometric morphometry of pottery shapes from Iron Age northeastern Taiwan

### 2020-01-27 analysis/paper/paper.Rmd
editing para on cv equality test on shapes

<!-- CV equality testing -->

Table \@ref(tab:metricCVstest) shows the results of Modified signed-likelihood ratio test (MSLRT) for equality of CVs for for six measurements and two ratio values. There is a significant difference (p-value less than 0.01) between CV values for two metric measurements across phases, the rim diameter and the neck diameter. The CV values show that the variation of rim diameter is small (6.8%)  before European contact, increases after the European presence (9.6%), and decreases after the Chinese presence (6.9%). For the neck diameter, the CV values indicate smaller variation after the European period (7.7%) and the Chinese period (7.1%) compared to pre-European period (10.6%). <!-- these numeric values in this paragraph should come from inline code, not copy-pasted in like this -->

@@

The variation of pottery shape for each phase is explored by the distributions of the first three PC scores to compare the extent of standardization across phases. Figure \@ref(fig:PCs-distribution) suggests variations of PC scores for each vessel by phase. The first PC, the hight of the vessels, shows more variations in the pre-European period compared to the Chinese period. That indicates standardization appears strong in the Chinese period but weak in the pre-European period. The second PC also presents a similar pattern that stronger standardization of shape in the Chinese period compared to the other two phases. To see whether the differences in the distribution of PCs between any two phases is statistically significant, coefficients of variation statistic (CVs) among phase groups of PCs is examined with a modified signed-likelihood ratio test for the equality of CVs. P-values for the significant test of CVs of PC1 and PC2 show significant differences (P-value less than 0.01) in the standardization of vessel shape across some periods, especially between Chinese contact with either pre-European or post-European(Figure \@ref(fig:CV-test-PCs-plot), Table \@ref(tab:CV-test-PCs-table)). Significant difference is also detected between pre-European and post-European (P-value less than 0.05).

>>>>

<!-- CV equality testing -->
To compare pottery shape standardization across the three phases we  investigate the distributions of the first three PC scores, taking the PC scores as proxy variables for vessel shape (Figure \@ref(fig:PCs-distribution)). The first PC, vessel height, shows higher variation in the pre-European period compared to the Chinese period. That indicates that shape standardization was higher in the Chinese period compared to the pre-European period. The second PC also presents a similar pattern of higher shape standardization in the Chinese period compared to the other two phases. To see whether the differences in the distribution of PCs between any two phases are substantive or due to chance, we used a modified signed-likelihood ratio test to assess the equality of CVs. P-values for this significance test of CVs for PC1 and PC2 show significant differences in the standardization of vessel shapes across periods, especially between Chinese contact with either pre-European or post-European(Figure \@ref(fig:CV-test-PCs-plot), Table \@ref(tab:CV-test-PCs-table)). A significant difference is also detected between the pre-European and post-European periods.

Table \@ref(tab:metricCVstest) shows the results of Modified signed-likelihood ratio test (MSLRT) for equality of CVs for for six measurements and two ratio values. There is a significant difference between CV values for two metric measurements across phases, the rim diameter and the neck diameter. The CV values show that the variation of rim diameter is small (6.8%)  before European contact, increases after the European presence (9.6%), and decreases after the Chinese presence (6.9%). For the neck diameter, the CV values indicate smaller variation after the European period (7.7%) and the Chinese period (7.1%) compared to pre-European period (10.6%). <!-- these numeric values in this paragraph should come from inline code, not copy-pasted in like this -->

@@
